---
title: "2019-person-EDA"
author: "Albert Sun"
date: "2/17/2022"
output: html_document
---

```{r include=FALSE}
library(tidyverse)
library(readr)
```

```{r include=FALSE}
df_raw <- read_csv('../../../data/17-19survey-person.csv')
glimpse(df_raw)
```

```{r}
household <- read_csv('../../../data/2017-2019-pr2-1-Household.csv')
glimpse(household)
```

```{r}
household%>%
  select(prev_home_rgcname) 
```

```{r}
household%>%
  select(final_home_rgcnum) 

```

```{r}
# queries for suburbanised people

stay_in_washington <- household %>%
  filter(sample_county == 'King') %>%
  filter(prev_home_wa == 'Yes, previous home was in Washington')

stay_in_washington
```

Here are data columns that are particularly interesting: 
- `final_home_tract` - census tract of where the final home is located
- `prev_home_tract` - census tract of where the first home is located

The breakdown for FIPS is 
- 2 - State
- 3 - County
- 6 - Tract
- 1 - Block Group

This website allows you to see the boundaries of specific census tracts: 

https://www.policymap.com/newmaps#/

This leads to the question - how do I categorize tracts into larger groups? 

- GIS of all census tracts in King County: https://www5.kingcounty.gov/sdc/Metadata.aspx?Layer=tracts10_shore


```{r}
stay_in_washington %>%
  select(hhid, final_home_tract, prev_home_tract, final_home_rgcnum, final_home_uvnum, prev_home_rgcname)
```

```{r}
df_prev_home <- stay_in_washington %>%
  group_by(prev_home_tract) %>%
  summarise(n = n()) 

df_prev_home
```

Top locations are: 
- Redmond, 
- Seattle West Edge/Central Business District, 
- Seattle Bell Town, 
- Capitol Hill, 
- Between University District/Wallingford

```{r}
df_final_home <- stay_in_washington %>%
  group_by(final_home_tract) %>%
  summarise(n = n())

df_final_home
```

Top locations are: 
- Redmond, 
- Northgate (Northern Seattle)
- Between University District/Wallingford
- Southlake Union (near airport)
- Capitol Hill

In order to track suburbanisation in this dataset, I think we would need a systematic way to categorize whether a specific census tract is suburban or not. Having a dataset that has a list of all the census tracts and what type of locations they describe would be a good start to doing this. 

```{r joining the tracts}
home_tracts <- full_join(df_prev_home, df_final_home, 
                         by=c('prev_home_tract' = 'final_home_tract')) %>%
  rename(tract = prev_home_tract,
         num_prev = n.x,
         num_final = n.y)

home_tracts

write_csv(home_tracts, "../../../data/survey-household-tracts-moving.csv")
```
